---
title: "Power Study Postprocessing and Localization"
author: "Alea Seifert"
format:
  html: 
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
---

## Setup

```{r setup}
library(dplyr)
library(purrr)
library(stringr)
library(ggplot2)
library(tidyr)
library(ggh4x)

OUT_DIR <- "C:/Users/joyse/OneDrive/Documents/OceanAcidification/lomad/dev/sims/power/results"
IMG_DIR <- file.path(OUT_DIR, "_img")
alpha   <- 0.05

dir.create(IMG_DIR, showWarnings = FALSE, recursive = TRUE)
```

## Load simulation-level results

```{r load-results}
results <- readRDS(file.path(OUT_DIR, "_simulations-power-results.rds"))
```

## Summarize detection probability

For each simulation condition, this computes the detection rate and an exact binomial confidence interval.

```{r summarize-results}
results_summary <- results |>
  group_by(struct, d, n, snr, phi, method) |>
  summarise(
    S         = n(),
    x         = sum(detected, na.rm = TRUE),
    detection = mean(detected, na.rm = TRUE),
    ci_lo     = if_else(x == 0, 0, qbeta(0.025, x, S - x + 1)),
    ci_hi     = if_else(x == S, 1, qbeta(0.975, x + 1, S - x)),
    .groups   = "drop"
  )

results_summary
```

## Power curves

```{r power-curves-all-n, fig.width=9, fig.height=5}
results_summary |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(d, detection, colour = struct,
             linetype = method,
             group = interaction(struct, method))) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             colour = "grey60", linewidth = 0.5) +
  geom_line(linewidth = 0.5, alpha = 0.8) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi, fill = struct),
              alpha = 0.2, colour = NA) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(values = c("Smooth" = "#0072B2",
                                 "Cross"  = "#D55E00",
                                 "Rate"   = "#009E73")) +
  scale_fill_manual(values = c("Smooth" = "#0072B2",
                               "Cross"  = "#D55E00",
                               "Rate"   = "#009E73")) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(x = "Separation (d)", y = "Power",
       colour = "Structure", fill = "Structure") +
  guides(linetype = guide_none()) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        axis.text = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.1, color = "darkgray"))

ggsave(file.path(IMG_DIR, "power_curves.png"), 
       width = 9, height = 4, dpi = 450)
```

## Localization helper functions

```{r localization-helpers}
RAW_DIR <- file.path(OUT_DIR, "_raw")

series_to_df <- function(rep_series) {
  vi <- rep_series$vi
  
  tibble(
    t = vi,
    sep = rep_series$sep[vi],
    w = if (!is.null(rep_series$w)) rep_series$w[vi] else rep(NA_real_, length(vi)),
    rejected = rep_series$rejected[vi]
  )
}

compute_sens_curve <- function(pool, col, c_grid) {
  vals <- pool[[col]]
  rej  <- pool$rejected
  vapply(c_grid, function(c) {
    above <- vals > c
    if (sum(above, na.rm = TRUE) == 0L) return(NA_real_)
    mean(rej[above], na.rm = TRUE)
  }, double(1))
}

compute_spec_curve <- function(pool, col, c_grid) {
  vals <- pool[[col]]
  rej  <- pool$rejected
  vapply(c_grid, function(c) {
    below <- vals <= c
    if (sum(below, na.rm = TRUE) == 0L) return(NA_real_)
    mean(!rej[below], na.rm = TRUE)
  }, double(1))
}

truth_given_class <- function(df) {
  df |>
    filter(!is.na(rejected)) |>
    group_by(rejected) |>
    summarise(
      n_windows  = n(),
      mean_sep   = mean(sep, na.rm = TRUE),
      median_sep = median(sep, na.rm = TRUE),
      q75_sep    = quantile(sep, 0.75, na.rm = TRUE),
      q90_sep    = quantile(sep, 0.90, na.rm = TRUE),
      max_sep    = max(sep, na.rm = TRUE),
      .groups    = "drop"
    )
}

parse_series_filename <- function(fname) {
  bn <- basename(fname) |> str_remove("-series\\.rds$")
  oracle <- str_detect(bn, "-oracle")
  bn <- str_remove(bn, "-oracle")
  parts <- str_match(
    bn,
    "^(\\w+)_d([0-9-]+)_n(\\d+)_snr([0-9-]+)_phi([0-9-]+)$"
  )
  tibble(
    file   = fname,
    struct = parts[, 2],
    d      = as.numeric(str_replace(parts[, 3], "-", ".")),
    n      = as.integer(parts[, 4]),
    snr    = as.numeric(str_replace(parts[, 5], "-", ".")),
    phi    = as.numeric(str_replace(parts[, 6], "-", ".")),
    oracle = oracle
  )
}

pool_reps <- function(series_obj) {
  imap_dfr(series_obj, function(rep, seed_name) {
    if (is.null(rep)) return(NULL)
    series_to_df(rep) |> mutate(seed = seed_name)
  })
}
```

## Load localization series files

```{r load-localization-files}
series_files <- list.files(RAW_DIR, pattern = "-series\\.rds$", full.names = TRUE)

file_meta <- map_dfr(series_files, parse_series_filename) |>
  filter(!is.na(d), d > 0, !oracle)
```

## Pilot localization settings

```{r pilot-localization-settings}
pick_file <- function(meta, struct_, d_, n_, snr_, phi_) {
  meta |>
    filter(struct == struct_, d == d_, n == n_, snr == snr_, phi == phi_) |>
    pull(file)
}


# Create h_win and s_win functions
h_win_fcn <- function(n) max(5L, floor(n / 200L))
s_win_fcn <- function(n) min(60L * h_win_fcn(n), floor(n / 4L))

# Choose settings once
fav_n <- 600
unfav_n <- 600

unfav_file <- pick_file(file_meta, "smooth", 1,   unfav_n, 1.5, 0.8)
fav_file   <- pick_file(file_meta, "smooth", 1.5, fav_n,   0.5, 0.3)

fav_series   <- readRDS(fav_file)
unfav_series <- readRDS(unfav_file)

pool_fav   <- pool_reps(fav_series)
pool_unfav <- pool_reps(unfav_series)

# Window sizes come from the n used above
fav_h_win   <- h_win_fcn(fav_n)
unfav_h_win <- h_win_fcn(unfav_n)
```

```{r}
make_random_facet_df <- function(series_obj,
                                 setting_name,
                                 n_seeds = 9,
                                 seed_value = 123) {
  set.seed(seed_value)
  
  chosen_seeds <- sample(names(series_obj), size = n_seeds)
  
  map_dfr(chosen_seeds, function(s) {
    series_to_df(series_obj[[s]]) |>
      mutate(
        seed = s,
        setting = setting_name
      )
  })
}
```

```{r}

fav_plot_df <- make_random_facet_df(
  fav_series,
  setting_name = "Favorable setting",
  n_seeds = 9,
  seed_value = 123
)

unfav_plot_df <- make_random_facet_df(
  unfav_series,
  setting_name = "Unfavorable setting",
  n_seeds = 9,
  seed_value = 456
)

fav_plot_df <- fav_plot_df |>
  group_by(seed) |>
  arrange(t) |>
  mutate(
    roll_sep = slide_dbl(
      sep,
      mean,
      .before = fav_h_win,
      na.rm = TRUE
    )) |>
  ungroup()

unfav_plot_df <- unfav_plot_df |>
  group_by(seed) |>
  arrange(t) |>
   mutate(
    roll_sep = slide_dbl(
      sep,
      mean,
      .before = unfav_h_win,
      na.rm = TRUE
    )) |>
  ungroup()
```


## Ground-truth separation by rejection class

```{r truth-given-class}
truth_given_class(pool_fav)
truth_given_class(pool_unfav)
```

## Sensitivity curve

This curve estimates

$$
P(\text{rejected} \mid sep_t > c).
$$

```{r sensitivity-curve, fig.width=8, fig.height=5}
c_grid <- seq(0, max(c(pool_fav$sep, pool_unfav$sep), na.rm = TRUE),
              length.out = 200)

sens_df <- bind_rows(
  tibble(c = c_grid,
         p_reject = compute_sens_curve(pool_fav, "sep", c_grid),
         setting  = "Unfavorable (structure=smooth, d=1, snr=1.5, phi=0.8)"),
  tibble(c = c_grid,
         p_reject = compute_sens_curve(pool_unfav, "sep", c_grid),
         setting  = "Favorable (structure=smooth, d=1, snr=0.5, phi=0.3)")
)

ggplot(sens_df, aes(x = c, y = p_reject, colour = setting)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = alpha, linetype = "dashed", colour = "grey60") +
  labs(x = expression("Threshold " * c),
       y = expression(P(rejected ~ "|" ~ sep[t] > c)),
       title = "Sensitivity curve — pointwise trend separation",
       colour = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

## Specificity curve

This curve estimates

$$
P(\text{not rejected} \mid sep_t \le c).
$$

```{r specificity-curve, fig.width=8, fig.height=5}
spec_df <- bind_rows(
  tibble(c = c_grid,
         p_notrej = compute_spec_curve(pool_fav, "sep", c_grid),
         setting  = "Unfavorable (structure=smooth, d=1, snr=1.5, phi=0.8)"),
  tibble(c = c_grid,
         p_notrej = compute_spec_curve(pool_unfav, "sep", c_grid),
         setting  = "Favorable (structure=smooth, d=1, snr=0.5, phi=0.3)")
)

ggplot(spec_df, aes(x = c, y = p_notrej, colour = setting)) +
  geom_line(linewidth = 0.8) +
  labs(x = expression("Threshold " * c),
       y = expression(P(not~rejected ~ "|" ~ sep[t] <= c)),
       title = "Specificity curve — pointwise trend separation",
       colour = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

## Individual replicate visualization

```{r individual-replicate, fig.width=8, fig.height=4}
rep_stats <- pool_fav |>
  group_by(seed) |>
  summarise(prop_rej = mean(rejected, na.rm = TRUE), .groups = "drop")

mixed_seed <- rep_stats |>
  arrange(abs(prop_rej - 0.5)) |>
  slice(1) |>
  pull(seed)

mixed_df <- pool_fav |> filter(seed == mixed_seed)

ggplot(mixed_df, aes(x = t, y = sep)) +
  geom_line(linewidth = 0.4) +
  geom_point(data = \(x) filter(x, rejected), size = 1.2, colour = "#D55E00") +
  labs(x = "Time index",
       y = expression("|" * x[1*t] - x[2*t] * "|"),
       title = "Localization: trend separation with rejected windows",
       subtitle = paste("seed", mixed_seed)) +
  theme_minimal(base_size = 13)
```

#### Favorable replicate visualization
```{r}
# Randomly select seeds
# 1. Favorable: separation with rejected points

ggplot(fav_plot_df, aes(x = t, y = sep)) +
  geom_line(linewidth = 0.4) +
  geom_point(
    data = \(x) filter(x, rejected),
    size = 0.8,
    colour = "#D55E00"
  ) +
  facet_wrap(~ seed, ncol = 3) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Favorable setting: separation over time",
    subtitle = "Points indicate rejected windows",
    x = "Time index",
    y = "Separation"
  )

# 2. Favorable: rolling seperation with rejected points

ggplot(fav_plot_df, aes(x = t, y = roll_sep)) +
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.3) +
  geom_line(linewidth = 0.4) +
  geom_point(
    data = \(x) filter(x, rejected),
    size = 0.8,
    colour = "#D55E00"
  ) +
  facet_wrap(~ seed, ncol = 3) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Favorable setting: rolling seperation over time",
    subtitle = "Rolling Seperation = average seperation of h_win points previous",
    x = "Time index",
    y = "Mean rolling seperation"
  )

```
```{r}
viz_structs <- c("smooth", "cross", "rate")
viz_d   <- 1
viz_n   <- 600
viz_snr <- 0.5
viz_phi <- 0.3

n_seeds_per_structure <- 3
seed_value <- 123

set.seed(seed_value)

series_by_struct <- set_names(viz_structs) |>
  map(function(struct_name) {
    file <- pick_file(
      file_meta,
      struct_name,
      viz_d,
      viz_n,
      viz_snr,
      viz_phi
    )

    readRDS(file)
  })

common_seeds <- Reduce(intersect, lapply(series_by_struct, names))

set.seed(seed_value)
chosen_seeds <- sample(common_seeds, n_seeds_per_structure)

all_struct_plot_df <- imap_dfr(
  series_by_struct,
  function(series_obj, struct_name) {

    map_dfr(chosen_seeds, function(s) {
      series_to_df(series_obj[[s]]) |>
        mutate(
          seed = s,
          struct = struct_name
        )
    })
  }
) |>
  group_by(struct, seed) |>
  arrange(t, .by_group = TRUE) |>
  mutate(
    roll_sep = slider::slide_dbl(
      sep,
      mean,
      .before = s_win_fcn(viz_n),
      .complete = FALSE,
      na.rm = TRUE
    )
  ) |>
  ungroup() |>
  mutate(
    seed = factor(seed, levels = chosen_seeds),
    struct = factor(struct, levels = viz_structs)
  )


ggplot(all_struct_plot_df, aes(x = t, y = roll_sep)) +
  geom_line(linewidth = 0.5, colour = "black") +
  geom_point(
    data = \(x) filter(x, rejected),
    aes(x = t, y = roll_sep),
    size = 1.2,
    colour = "orange3"
  ) +
  facet_grid(
    rows = vars(struct),
    cols = vars(seed)
  ) +

  theme_minimal(base_size = 10) +
  labs(
    title = "Rolling mean separation across structures",
    subtitle = paste0(
      "Shared seeds; d = ", viz_d,
      ", n = ", viz_n,
      ", SNR = ", viz_snr,
      ", phi = ", viz_phi,
      ", s_win = ", s_win_fcn(viz_n)
    ),
    x = "Time index",
    y = "Rolling mean separation"
  )
```



#### Unfavorable replicate visualization
```{r}
# 1. Unfavorable: separation with rejected points

ggplot(unfav_plot_df, aes(x = t, y = sep)) +
  geom_line(linewidth = 0.4) +
  geom_point(
    data = \(x) filter(x, rejected),
    size = 0.8,
    colour = "#D55E00"
  ) +
  facet_wrap(~ seed, ncol = 3) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Unfavorable setting: separation over time",
    subtitle = "Points indicate rejected windows",
    x = "Time index",
    y = "Separation"
  )

# 2. Unfavorable: rolling seperation with rejected points

ggplot(unfav_plot_df, aes(x = t, y = roll_sep)) +
  geom_line(linewidth = 0.4) +
  geom_point(
    data = \(x) filter(x, rejected),
    size = 0.8,
    colour = "#D55E00"
  ) +
  facet_wrap(~ seed, ncol = 3) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Unfavorable setting: rolling seperation over time",
    subtitle = "Rolling Seperation = average seperation of h_win points previous",
    x = "Time index",
    y = "Mean rolling seperation"
  )
```

```{r}
viz_structs <- c("smooth", "cross", "rate")
viz_d   <- 2
viz_n   <- 600
viz_snr <- 1.5
viz_phi <- 0.8

n_seeds_per_structure <- 3
seed_value <- 123

set.seed(seed_value)

all_struct_plot_df <- map_dfr(viz_structs, function(struct_name) {
  
  file <- pick_file(file_meta, struct_name, viz_d, viz_n, viz_snr, viz_phi)
  series_obj <- readRDS(file)
  
  chosen_seeds <- sample(names(series_obj), size = n_seeds_per_structure)
  
  map_dfr(chosen_seeds, function(s) {
    series_to_df(series_obj[[s]]) |>
      mutate(
        seed = s,
        struct = struct_name,
        rep_id = paste(struct_name, s, sep = " | ")
      )
  })
}) |>
  group_by(struct, seed) |>
  arrange(t, .by_group = TRUE) |>
  mutate(
    roll_sep = slider::slide_dbl(
      sep,
      mean,
      .before = h_win_fcn(viz_n),
      .complete = FALSE,
      na.rm = TRUE
    )
  ) |>
  ungroup()

ggplot(all_struct_plot_df, aes(x = t, y = roll_sep)) +
  geom_line(linewidth = 0.5, colour = "black") +
  geom_point(
    data = \(x) filter(x, rejected),
    aes(x = t, y = roll_sep),
    size = 1.2,
    colour = "#D55E00"
  ) +
  facet_wrap(~ rep_id, ncol = 3) +
  theme_minimal(base_size = 13) +
  labs(
    title = "Rolling mean separation across structures",
    subtitle = paste0(
      "Random seeds; d = ", viz_d,
      ", n = ", viz_n,
      ", SNR = ", viz_snr,
      ", phi = ", viz_phi,
      ", s_win = ", s_win_fcn(viz_n)
    ),
    x = "Time index",
    y = "Rolling mean separation"
  )
```



#### Series Structures: fav vs. unfav
```{r}
## Load series files for all three structures (favorable setting)
fav_smooth <- readRDS(
  paste0(
    OUT_DIR,
    "/_raw/smooth_d1-0_n600_snr0-5_phi0-3-series.rds"
  )
)
fav_cross <- readRDS(
  paste0(
    OUT_DIR,
    "/_raw/cross_d1-0_n600_snr0-5_phi0-3-series.rds"
  )
)
fav_rate <- readRDS(
  paste0(
    OUT_DIR,
    "/_raw/rate_d1-0_n600_snr0-5_phi0-3-series.rds"
  )
)

# Pick 3 replicates from each that have some rejections
pick_reps <- function(series_obj, n_pick = 5) {
  has_rejection <- sapply(series_obj, function(rep) any(rep$rejected, na.rm = TRUE))
  chosen <- sample(names(series_obj)[has_rejection], n_pick)
  series_obj[chosen]
}


reps_smooth <- pick_reps(fav_smooth)
reps_cross  <- pick_reps(fav_cross)
reps_rate   <- pick_reps(fav_rate)

make_rep_df <- function(reps, structure) {
  lapply(seq_along(reps), function(i) {
    rep  <- reps[[i]]
    data.frame(
      t         = seq_along(rep$w),
      w         = rep$w,
      sep       = rep$sep,
      decoupling = abs(rep$sep),
      divergence = c(NA, abs(diff(rep$sep))),
      rejected  = rep$rejected,
      rep_id    = paste0(structure, " — rep ", i),
      structure = structure
    )
  }) |> bind_rows()
}

plot_df <- bind_rows(
  make_rep_df(reps_smooth, "smooth"),
  make_rep_df(reps_cross,  "cross"),
  make_rep_df(reps_rate,   "rate")
)

ggplot(plot_df, aes(x = t)) +
  geom_tile(aes(y = 0.1, height = Inf, fill = rejected), alpha = 0.3) +
  geom_line(aes(y = decoupling)) +
  scale_fill_manual(values = c("TRUE" = "#D55E00", "FALSE" = "grey90"),
                    na.value = "white") +
  facet_wrap(~ rep_id, ncol = 5) +
  labs(x = "Time", y =  expression("|" * x[1*t] - x[2*t] * "|"), fill = "Rejected",
       title = "Decoupling over time with rejected regions — All structure favorable setting") +
  theme_minimal() +
  theme(legend.position = "bottom")


ggplot(plot_df, aes(x = t)) +
  geom_tile(aes(y = 0.025, height = Inf, fill = rejected), alpha = 0.3) +
  geom_line(aes(y = divergence)) +
  scale_fill_manual(values = c("TRUE" = "#D55E00", "FALSE" = "grey90"),
                    na.value = "white") +
  facet_wrap(~ rep_id, ncol = 5) +
  labs(x = "Time", y =  expression("|" * sep[t] - sep[t-1] * "|"), fill = "Rejected",
       title = "Rate of Seperation over time with rejected regions — All structure favorable setting") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
## Load series files for all three structures (favorable setting)
unfav_smooth <- readRDS(
  paste0(
    OUT_DIR,
    "/_raw/smooth_d1-0_n600_snr1-5_phi0-8-series.rds"
  )
)
unfav_cross <- readRDS(
  paste0(
    OUT_DIR,
    "/_raw/cross_d1-0_n600_snr1-5_phi0-8-series.rds"
  )
)
unfav_rate <- readRDS(
  paste0(
    OUT_DIR,
    "/_raw/rate_d1-0_n600_snr1-5_phi0-8-series.rds"
  )
)

# Pick 3 replicates from each that have some rejections
pick_reps <- function(series_obj, n_pick = 5) {
  has_rejection <- sapply(series_obj, function(rep) any(rep$rejected, na.rm = TRUE))
  chosen <- sample(names(series_obj)[has_rejection], n_pick)
  series_obj[chosen]
}

set.seed(42)
reps_smooth <- pick_reps(unfav_smooth)
reps_cross  <- pick_reps(unfav_cross)
reps_rate   <- pick_reps(unfav_rate)

make_rep_df <- function(reps, structure) {
  lapply(seq_along(reps), function(i) {
    rep  <- reps[[i]]
    data.frame(
      t         = seq_along(rep$w),
      w         = rep$w,
      sep       = rep$sep,
      decoupling = abs(rep$sep),
      divergence = c(NA, abs(diff(rep$sep))),
      rejected  = rep$rejected,
      rep_id    = paste0(structure, " — rep ", i),
      structure = structure
    )
  }) |> bind_rows()
}

plot_df <- bind_rows(
  make_rep_df(reps_smooth, "smooth"),
  make_rep_df(reps_cross,  "cross"),
  make_rep_df(reps_rate,   "rate")
)

ggplot(plot_df, aes(x = t)) +
  geom_tile(aes(y = 0.1, height = Inf, fill = rejected), alpha = 0.3) +
  geom_line(aes(y = decoupling)) +
  scale_fill_manual(values = c("TRUE" = "#D55E00", "FALSE" = "grey90"),
                    na.value = "white") +
  facet_wrap(~ rep_id, ncol = 5) +
  labs(x = "Time", y =  expression("|" * x[1*t] - x[2*t] * "|"), fill = "Rejected",
       title = "Decoupling over time with rejected regions — All structure unfavorable setting") +
  theme_minimal() +
  theme(legend.position = "bottom")


ggplot(plot_df, aes(x = t)) +
  geom_tile(aes(y = 0.025, height = Inf, fill = rejected), alpha = 0.3) +
  geom_line(aes(y = divergence)) +
  scale_fill_manual(values = c("TRUE" = "#D55E00", "FALSE" = "grey90"),
                    na.value = "white") +
  facet_wrap(~ rep_id, ncol = 5) +
  labs(x = "Time", y =  expression("|" * sep[t] - sep[t-1] * "|"), fill = "Rejected",
       title = "Rate of Seperation over time with rejected regions — All structure unfavorable setting") +
  theme_minimal() +
  theme(legend.position = "bottom")
```


## Fixed-cutoff localization summaries

```{r fixed-cutoff-data}
# Load and pool all series files
file_meta <- file_meta[-69, ]

all_pools <- file_meta |>
  mutate(pool = map(file, \(f) readRDS(f) |> pool_reps())) |>
  unnest(pool) |>
  filter(!is.na(rejected)) |>
  mutate(method = if_else(oracle, "oracle", "estimated"))



# Specify cutoff threshold c
C_CUTOFF <- 0.07

# Specificity summary
spec_summary <- all_pools |>
  filter(sep <= C_CUTOFF) |>
  group_by(struct, d, n, snr, phi, method) |>
  summarise(
    S        = n(),
    n_notrej = sum(!rejected),
    spec     = mean(!rejected),
    spec_lo  = qbeta(0.025, n_notrej,     S - n_notrej + 1),
    spec_hi  = qbeta(0.975, n_notrej + 1, S - n_notrej),
    .groups  = "drop"
  )

# Sensitivity summary
sens_summary <- all_pools |>
  filter(sep > C_CUTOFF) |>
  group_by(struct, d, n, snr, phi, method) |>
  summarise(
    S       = n(),
    n_rej   = sum(rejected),
    sens    = mean(rejected),
    sens_lo = qbeta(0.025, n_rej,     S - n_rej + 1),
    sens_hi = qbeta(0.975, n_rej + 1, S - n_rej),
    .groups = "drop"
  )

```

#### Fixed-cutoff specificity plot

```{r fixed-cutoff-specificity, fig.width=9, fig.height=5}
spec_summary |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(d, spec, colour = struct,
             linetype = method,
             group = interaction(struct, method))) +
  geom_hline(yintercept = 1 - alpha, linetype = "dashed",
             colour = "grey60", linewidth = 0.5) +
  geom_line(linewidth = 0.5, alpha = 0.8) +
  geom_ribbon(aes(ymin = spec_lo, ymax = spec_hi, fill = struct),
              alpha = 0.2, colour = NA) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(values = c("Smooth" = "#0072B2",
                                 "Cross"  = "#D55E00",
                                 "Rate"   = "#009E73")) +
  scale_fill_manual(values = c("Smooth" = "#0072B2",
                               "Cross"  = "#D55E00",
                               "Rate"   = "#009E73")) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(x      = "Separation (d)",
       y      = paste0("Specificity (c = ", C_CUTOFF, ")"),
       colour = "Structure",
       fill   = "Structure",
       caption = paste0("Specificity = P(not rejected | sep <= ", C_CUTOFF, ")")) +
  guides(linetype = guide_none()) +
  theme_minimal(base_size = 12) +1
  theme(legend.position  = "right",
        axis.text        = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.1, color = "darkgray"))

ggsave(file.path(IMG_DIR, "spec_curves.png"),
       width = 9, height = 4, dpi = 450)
```

#### Fixed-cutoff sensitivity plot

```{r fixed-cutoff-sensitivity, fig.width=9, fig.height=5}
sens_summary |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(d, sens, colour = struct,
             linetype = method,
             group = interaction(struct, method))) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             colour = "grey60", linewidth = 0.5) +
  geom_line(linewidth = 0.5, alpha = 0.8) +
  geom_ribbon(aes(ymin = sens_lo, ymax = sens_hi, fill = struct),
              alpha = 0.2, colour = NA) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(values = c("Smooth" = "#0072B2",
                                 "Cross"  = "#D55E00",
                                 "Rate"   = "#009E73")) +
  scale_fill_manual(values = c("Smooth" = "#0072B2",
                               "Cross"  = "#D55E00",
                               "Rate"   = "#009E73")) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(x      = "Separation (d)",
       y      = paste0("Sensitivity (c = ", C_CUTOFF, ")"),
       colour = "Structure",
       fill   = "Structure", 
       caption = paste0("Sensitivity = P(rejected | sep > ", C_CUTOFF, ")")) +
  guides(linetype = guide_none()) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "right",
        axis.text        = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.1, color = "darkgray"))

```
## Double encoded c localization 
```{r}
C_CUTOFFS <- c(0.02, 0.08)

# Specificity summary

spec_summary_multi <- map_dfr(C_CUTOFFS, function(cutoff) {
  all_pools |>
    filter(sep <= cutoff) |>
    group_by(struct, d, n, snr, phi, method) |>
    summarise(
      S        = n(),
      n_notrej = sum(!rejected),
      spec     = mean(!rejected),
      spec_lo  = if_else(n_notrej == 0, 0, qbeta(0.025, n_notrej, S - n_notrej + 1)),
      spec_hi  = if_else(n_notrej == S, 1, qbeta(0.975, n_notrej + 1, S - n_notrej)),
      .groups  = "drop"
    ) |>
    mutate(cutoff = paste0("c = ", cutoff))
})


sens_summary_multi <- map_dfr(C_CUTOFFS, function(cutoff) {
  all_pools |>
    filter(sep > cutoff) |>
    group_by(struct, d, n, snr, phi, method) |>
    summarise(
      S       = n(),
      n_rej   = sum(rejected),
      sens    = mean(rejected),
      sens_lo = if_else(n_rej == 0, 0, qbeta(0.025, n_rej, S - n_rej + 1)),
      sens_hi = if_else(n_rej == S, 1, qbeta(0.975, n_rej + 1, S - n_rej)),
      .groups = "drop"
    ) |>
    mutate(cutoff = paste0("c = ", cutoff))
})
```

#### Sepecificity
```{r}
spec_summary_multi |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(
    x = d,
    y = spec,
    colour = struct,
    linetype = cutoff,
    group = interaction(struct, method, cutoff)
  )) +
  geom_hline(
    yintercept = 1 - alpha,
    linetype = "dashed",
    colour = "grey60",
    linewidth = 0.5
  ) +
  geom_line(linewidth = 0.5, alpha = 0.9) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(
    values = c(
      "Smooth" = "#0072B2",
      "Cross"  = "#D55E00",
      "Rate"   = "#009E73"
    )
  ) +
  scale_linetype_manual(
    values = c(
      "c = 0.02" = "solid",
      "c = 0.08" = "dashed"
    )
  ) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(
    x = "Separation (d)",
    y = "Specificity",
    colour = "Structure",
    linetype = "Threshold",
    caption = "Specificity = P(not rejected | sep <= c)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position  = "right",
    axis.text        = element_text(size = 8),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(linewidth = 0.1, color = "darkgray")
  )
```
#### Sensitivity
```{r}
sens_summary_multi |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(
    x = d,
    y = sens,
    colour = struct,
    linetype = cutoff,
    group = interaction(struct, method, cutoff)
  )) +
  geom_hline(
    yintercept = 1 - alpha,
    linetype = "dashed",
    colour = "grey60",
    linewidth = 0.5
  ) +
  geom_line(linewidth = 0.5, alpha = 0.9) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(
    values = c(
      "Smooth" = "#0072B2",
      "Cross"  = "#D55E00",
      "Rate"   = "#009E73"
    )
  ) +
  scale_linetype_manual(
    values = c(
      "c = 0.02" = "solid",
      "c = 0.08" = "dashed"
    )
  ) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(
    x = "Separation (d)",
    y = "Sensitivity",
    colour = "Structure",
    linetype = "Threshold",
    caption = "Sensitivity = P(rejected | sep > c)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position  = "right",
    axis.text        = element_text(size = 8),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(linewidth = 0.1, color = "darkgray")
  )
```
## Rolling window fixed c cutoff
```{r}
### played around with cutoff
C_CUTOFF <- 0.05

# Create h_win and s_win functions
h_win_fcn <- function(n) max(5L, floor(n / 200L))
s_win_fcn <- function(n) min(60L * h_win_fcn(n), floor(n / 4L))

# Create rolling window over sep
# Aggregate sep over [t - s_win + 1, t] 
# Make modular => can choose max or average over window
# agg_fcn = mean or max
windowed_sep <- function(sep, t, s_win, agg_fcn = mean) {
  slider::slide_index_dbl(
    sep, t, \(x) agg_fcn(x, na.rm = TRUE),
    .before   = s_win - 1L,
    .complete = FALSE
  )
}

# Compute sep_win for each series
# Nest by n so s_win derived correctly for each series length
# Group by condition combos
all_pools_w <- all_pools |>
  nest(.by = n) |>
  mutate(data = map2(data, n, \(df, n_val) {
    s_win <- s_win_fcn(n_val)
    df |>
      group_by(struct, d, snr, phi, method, seed) |>
      arrange(t, .by_group = TRUE) |>
      mutate(sep_win = windowed_sep(sep, t, s_win, agg_fcn = max)) |>  # swap max <-> mean here
      ungroup()
  })) |>
  unnest(data)
```


#### Specificity
```{r}
# Specificity summary (windowed)
win_spec_summary <- all_pools_w |>
  filter(sep_win <= C_CUTOFF) |>
  group_by(struct, d, n, snr, phi, method) |>
  summarise(
    S        = n(),
    n_notrej = sum(!rejected),
    spec     = mean(!rejected),
    spec_lo  = qbeta(0.025, n_notrej,     S - n_notrej + 1),
    spec_hi  = qbeta(0.975, n_notrej + 1, S - n_notrej),
    .groups  = "drop"
  ) |>
  summarise(
    S       = n(),
    n_rej   = sum(rejected),
    sens    = mean(rejected),
    sens_lo = qbeta(0.025, n_rej,     S - n_rej + 1),
    sens_hi = qbeta(0.975, n_rej + 1, S - n_rej),
    .groups = "drop"
  )



## Windowed specificity plot ####
win_spec_summary |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(d, spec, colour = struct,
             linetype = method,
             group = interaction(struct, method))) +
  geom_hline(yintercept = 1 - alpha, linetype = "dashed",
             colour = "grey60", linewidth = 0.5) +
  geom_line(linewidth = 0.5, alpha = 0.8) +
  geom_ribbon(aes(ymin = spec_lo, ymax = spec_hi, fill = struct),
              alpha = 0.2, colour = NA) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(values = c("Smooth" = "#0072B2",
                                 "Cross"  = "#D55E00",
                                 "Rate"   = "#009E73")) +
  scale_fill_manual(values = c("Smooth" = "#0072B2",
                               "Cross"  = "#D55E00",
                               "Rate"   = "#009E73")) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(x      = "Separation (d)",
       y      = paste0("Specificity = P(not rejected | win_sep \u2264 ", C_CUTOFF, ")"),
       colour = "Structure",
       fill   = "Structure") +
  guides(linetype = guide_none()) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "right",
        axis.text        = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.1, color = "darkgray"))
```



#### Sensitivity
```{r}
# Sensitivity summary (windowed)
win_sens_summary <- all_pools_w |>
  filter(sep_win > C_CUTOFF) |>
  group_by(struct, d, n, snr, phi, method) |>
  summarise(
    S       = n(),
    n_rej   = sum(rejected),
    sens    = mean(rejected),
    sens_lo = qbeta(0.025, n_rej,     S - n_rej + 1),
    sens_hi = qbeta(0.975, n_rej + 1, S - n_rej),
    .groups = "drop"
  )


## Windowed sensitivity plot ####
win_sens_summary |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(d, sens, colour = struct,
             linetype = method,
             group = interaction(struct, method))) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             colour = "grey60", linewidth = 0.5) +
  geom_line(linewidth = 0.5, alpha = 0.8) +
  geom_ribbon(aes(ymin = sens_lo, ymax = sens_hi, fill = struct),
              alpha = 0.2, colour = NA) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(values = c("Smooth" = "#0072B2",
                                 "Cross"  = "#D55E00",
                                 "Rate"   = "#009E73")) +
  scale_fill_manual(values = c("Smooth" = "#0072B2",
                               "Cross"  = "#D55E00",
                               "Rate"   = "#009E73")) +
  facet_nested(phi ~ snr + n, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    n   = \(x) paste0("T == ", x),
    .default = label_parsed
  )) +
  labs(x      = "Separation (d)",
       y      = paste0("Sensitivity = P(rejected | win_sep > ", C_CUTOFF, ")"),
       colour = "Structure",
       fill   = "Structure") +
  guides(linetype = guide_none()) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "right",
        axis.text        = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.1, color = "darkgray"))

```


## ROC
- sensitivity x 
- 1- specificity y 
- roc with line structure
- s-win top
- snr z still
- c grid, then summarixe and calculate both 
- pool all data and then specifiy 
= data frame: c, (SNR phi, structure fixed), sensitivity, and specificity ...
```{r}
# pool data
OUT_DIR <- "C:/Users/joyse/OneDrive/Documents/OceanAcidification/lomad/dev/sims/power/results"
RAW_DIR <- file.path(OUT_DIR, "_raw")
series_files <- list.files(RAW_DIR, pattern = "-series\\.rds$", full.names = TRUE)

series_to_df <- function(rep_series) {
  vi <- rep_series$vi
  
  tibble(
    t = vi,
    sep = rep_series$sep[vi],
    w = if (!is.null(rep_series$w)) rep_series$w[vi] else rep(NA_real_, length(vi)),
    rejected = rep_series$rejected[vi]
  )
}

parse_series_filename <- function(fname) {
  bn <- basename(fname) |> str_remove("-series\\.rds$")
  oracle <- str_detect(bn, "-oracle")
  bn <- str_remove(bn, "-oracle")
  parts <- str_match(
    bn,
    "^(\\w+)_d([0-9-]+)_n(\\d+)_snr([0-9-]+)_phi([0-9-]+)$"
  )
  tibble(
    file   = fname,
    struct = parts[, 2],
    d      = as.numeric(str_replace(parts[, 3], "-", ".")),
    n      = as.integer(parts[, 4]),
    snr    = as.numeric(str_replace(parts[, 5], "-", ".")),
    phi    = as.numeric(str_replace(parts[, 6], "-", ".")),
    oracle = oracle
  )
}

pool_reps <- function(series_obj) {
  imap_dfr(series_obj, function(rep, seed_name) {
    if (is.null(rep)) return(NULL)
    series_to_df(rep) |> mutate(seed = seed_name)
  })
}

file_meta <- map_dfr(series_files, parse_series_filename) |>
  filter(!is.na(d), d > 0, !oracle)

file_meta <- file_meta[-69, ]


all_pools <- file_meta |>
  mutate(pool = map(file, \(f) readRDS(f) |> pool_reps())) |>
  unnest(pool) |>
  filter(!is.na(rejected)) |>
  mutate(method = if_else(oracle, "oracle", "estimated"))

######################################################################################################################

# Create h_win and s_win functions
h_win_fcn <- function(n) max(5L, floor(n / 200L))
s_win_fcn <- function(n) min(60L * h_win_fcn(n), floor(n / 4L))

# Create rolling window over sep
# Aggregate sep over [t - s_win + 1, t] 
# Make modular => can choose max or average over window
# agg_fcn = mean or max
windowed_sep <- function(sep, t, s_win, agg_fcn = mean) {
  slider::slide_index_dbl(
    sep, t, \(x) agg_fcn(x, na.rm = TRUE),
    .before   = s_win - 1L,
    .complete = FALSE
  )
}

# find max and mean sep of s_win
all_pools_w <- all_pools |>
  nest(.by = n) |>
  mutate(data = map2(data, n, \(df, n_val) {
    s_win_val <- s_win_fcn(n_val)

    df |>
      group_by(struct, d, snr, phi, method, seed) |>
      arrange(t, .by_group = TRUE) |>
      mutate(
        s_win = s_win_val,
        sep_val_mean = windowed_sep(sep, t, s_win_val, agg_fcn = mean),
        sep_val_max  = windowed_sep(sep, t, s_win_val, agg_fcn = max)
      ) |>
      ungroup()
  })) |>
  unnest(data)

# Create c grid
c_grid <- seq(0, 0.3, length.out = 100)

# specify distance
all_pools_d <- all_pools_w |>
  select(struct, snr, phi, sep, d, s_win, rejected, sep_val_mean, sep_val_max)
```

```{r}
###################### summary (windowed) mean
metric_summary_mean <- map_dfr(c_grid, function(cutoff) {
  all_pools_d |>
    mutate(pred_reject = sep_val_mean > cutoff) |>
    group_by(struct, snr, phi, s_win) |>
    summarise(
      c = cutoff,
      tp = sum(pred_reject & rejected, na.rm = TRUE),
      fn = sum(!pred_reject & rejected, na.rm = TRUE),
      tn = sum(!pred_reject & !rejected, na.rm = TRUE),
      fp = sum(pred_reject & !rejected, na.rm = TRUE),
      sens = tp / (tp + fn),
      spec = tn / (tn + fp),
      fpr = 1 - spec,
      .groups = "drop"
    )
})


metric_summary_max <- map_dfr(c_grid, function(cutoff) {
  all_pools_d |>
    mutate(pred_reject = sep_val_max > cutoff) |>
    group_by(struct, snr, phi, s_win) |>
    summarise(
      c = cutoff,
      tp = sum(pred_reject & rejected, na.rm = TRUE),
      fn = sum(!pred_reject & rejected, na.rm = TRUE),
      tn = sum(!pred_reject & !rejected, na.rm = TRUE),
      fp = sum(pred_reject & !rejected, na.rm = TRUE),
      sens = tp / (tp + fn),
      spec = tn / (tn + fp),
      
      fpr = 1 - spec,
      .groups = "drop"
    )
})


## Windowed AUC plot ####
metric_summary_mean |>
arrange(struct, phi, snr, s_win, c) |>
  mutate(
    struct = str_to_sentence(struct)
  ) |>
  ggplot(aes(
    x = fpr,
    y = sens,
    color = struct,
    group = interaction(struct, phi, snr, s_win)
  )) +
  geom_path(linewidth = 0.5, alpha = 0.8) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed",
              colour = "grey60") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  facet_nested(phi ~ snr + s_win, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    s_win = \(x) paste0("s == ", x),
    .default = label_parsed
  )) +
  labs(
    title = "ROC: Mean seperation of s window",
    x = "1 - specificity",
    y = "Sensitivity",
    colour = "Structure"
  ) +
  theme_minimal(base_size = 12)


## Windowed AUC plot ####
metric_summary_max |>
arrange(struct, phi, snr, s_win, c) |>
  mutate(
    struct = str_to_sentence(struct)
  ) |>
  ggplot(aes(
    x = fpr,
    y = sens,
    color = struct,
    group = interaction(struct, phi, snr, s_win)
  )) +
  geom_path(linewidth = 0.5, alpha = 0.8) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed",
              colour = "grey60") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  facet_nested(phi ~ snr + s_win, labeller = labeller(
    phi = \(x) paste0("phi == ", x),
    snr = \(x) paste0("SNR == ", x),
    s_win = \(x) paste0("s == ", x),
    .default = label_parsed
  )) +
  labs(
    title = "ROC: Max seperation of s window",
    x = "1 - specificity",
    y = "Sensitivity",
    colour = "Structure"
  ) +
  theme_minimal(base_size = 12)
```


```{r}
library(pracma)

auc_summary <- metric_summary_max |>
  group_by(struct, snr, phi, s_win) |>
  arrange(fpr, .by_group = TRUE) |>
  summarise(
    auc = trapz(fpr, sens),
    .groups = "drop"
  )

auc_summary |>
  mutate(struct = str_to_sentence(struct)) |>
  ggplot(aes(
    x = s_win,
    y = auc,
    color = struct,
    group = struct
  )) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 2) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_manual(values = c(
    "Smooth" = "#0072B2",
    "Cross"  = "#D55E00",
    "Rate"   = "#009E73"
  )) +
  facet_grid(phi ~ snr, labeller = labeller(
    phi = \(x) paste0("phi = ", x),
    snr = \(x) paste0("SNR = ", x)
  )) +
  labs(
    title = "Area under curve",
    x = "Rolling window size",
    y = "AUC",
    colour = "Structure"
  ) +
  theme_minimal(base_size = 13)
```


