---
title: "Simulating Coupled Trend Pairs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating Coupled Trend Pairs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

The `lomad` package provides `sim_trends()` for simulating pairs of smooth
trend series with controlled coupling behavior. These are useful for
benchmarking and evaluating decoupling detection methods under known
ground-truth conditions.

The primary parameter across all methods is `d`, the Euclidean distance
between the two output series. The `method` argument (or a custom `w` vector)
controls the *structure* of the coupling process (e.g. how often and how long
the series diverge). In the examples below, method-specific parameters are
held fixed so that the effect of `d` can be seen in isolation.

```{r libs}
library(lomad)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
```

## Shared helpers

`plot_dist()` is used for `method = "dist"`, which produces a constant
coupling weight (`w = 0`). `plot_trends()` is used for methods that return a
time-varying `w` — the shaded band marks the coupled region (w ∈ [0.5, 1.5]).

```{r helpers}
plot_trends <- function(sim, title = NULL) {
  n  <- length(sim$x1)
  df <- data.frame(t = seq_len(n), x1 = sim$x1, x2 = sim$x2,
                   x_mean = sim$x_mean, w = sim$w)

  p_series <- df |>
    pivot_longer(c(x1, x2, x_mean), names_to = "series", values_to = "value") |>
    ggplot(aes(x = t, y = value, color = series)) +
    geom_line(linewidth = 0.45) +
    scale_color_manual(values = c(x1 = "steelblue", x2 = "tomato",
                                  x_mean = "gray60")) +
    labs(title = title, y = NULL, color = NULL) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "none", axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          plot.title = element_text(size = 9))

  p_w <- ggplot(df, aes(x = t, y = w)) +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = 0.5, ymax = 1.5,
             fill = "gray90", alpha = 0.6) +
    geom_hline(yintercept = c(0, 1), linetype = "dashed",
               color = "gray70", linewidth = 0.3) +
    geom_line(linewidth = 0.4) +
    labs(y = "w", x = "t") +
    theme_minimal(base_size = 10)

  p_series / p_w + plot_layout(heights = c(2, 1))
}

plot_dist <- function(sim, title = NULL) {
  n <- length(sim$x1)
  data.frame(t = seq_len(n), x1 = sim$x1, x2 = sim$x2) |>
    pivot_longer(c(x1, x2), names_to = "series", values_to = "value") |>
    ggplot(aes(x = t, y = value, color = series)) +
    geom_line(linewidth = 0.45) +
    scale_color_manual(values = c(x1 = "steelblue", x2 = "tomato")) +
    labs(title = title, y = NULL, x = "t", color = NULL) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "none", plot.title = element_text(size = 9))
}

ds <- c(0.5, 1, 2, 4)
```

## `method = "dist"`

Generates two stationary trend series at a controlled coefficient-space
distance `d`. There is no coupling process — the series are independent
draws from a shared Fourier basis. Larger `d` produces more distinct
series with lower correlation.

```{r dist, fig.cap = "Two series at varying d."}
plots_dist <- lapply(ds, \(d)
  plot_dist(sim_trends(n = 500, d = d, seed = 7),
            title = sprintf("d = %.1f", d))
)

wrap_plots(plots_dist, ncol = 2) +
  plot_annotation(title = "method = 'dist': varying d")
```

## `method = "rate"`

Generates two trend series that repel symmetrically at periodic decoupling
events. `d` controls the Euclidean distance between the output series;
`rate` (events per unit time) controls how often events occur. The
`w` panel shows the structured, evenly-spaced dips.

```{r rate, fig.height = 8, fig.cap = "Periodic repulsion at varying d (rate = 0.01)."}
plots_rate <- lapply(ds, \(d)
  plot_trends(sim_trends(n = 500, d = d, method = "rate",
                          rate = 0.01, seed = 7),
              title = sprintf("d = %.1f", d))
)

wrap_plots(plots_rate, ncol = 2) +
  plot_annotation(title = "method = 'rate': varying d  (rate = 0.01)")
```

## `method = "smooth"`

Generates two trend series with organic, aperiodic repulsion governed by a
smooth stochastic `w` derived from kernel-smoothed Gaussian noise.
`w` is restricted to (0, 1) so the series always stay on their respective
sides of the mean. `bw` controls the timescale of coupling/decoupling
stretches; `coupling` controls the fraction of time the series track closely.

```{r smooth, fig.height = 8, fig.cap = "Stochastic repulsion at varying d (bw = 50, coupling = 0.8)."}
plots_smooth <- lapply(ds, \(d)
  plot_trends(sim_trends(n = 500, d = d, method = "smooth",
                          bw = 50, coupling = 0.8, seed = 7),
              title = sprintf("d = %.1f", d))
)

wrap_plots(plots_smooth, ncol = 2) +
  plot_annotation(title = "method = 'smooth': varying d  (bw = 50, coupling = 0.8)")
```

## `method = "cross"`

A variant of `method = "smooth"` where `w` is generated as a linear
process centered at 1, allowing it to exceed 1 (crossing: the series swap
sides of the mean) as well as drop below 0 (strong decoupling). The
`coupling` parameter controls the fraction of time `w` stays within 0.5
of 1, with the remainder split equally between decoupling and crossing tails.

```{r cross, fig.height = 8, fig.cap = "Stochastic coupling with crossings at varying d (bw = 50, coupling = 0.8)."}
plots_cross <- lapply(ds, \(d)
  plot_trends(sim_trends(n = 500, d = d, method = "cross",
                          bw = 50, coupling = 0.8, seed = 7),
              title = sprintf("d = %.1f", d))
)

wrap_plots(plots_cross, ncol = 2) +
  plot_annotation(title = "method = 'cross': varying d  (bw = 50, coupling = 0.8)")
```

`w` values above the shaded band (w > 1.5) indicate crossing events,
where the two series have swapped sides of their shared mean.

## Summary

| Method | Coupling structure | w range | Key parameters |
|---|---|---|---|
| `"dist"` | None (static) | 0 | `d` |
| `"rate"` | Periodic gamma bumps | [0, 1] | `d`, `rate` |
| `"smooth"` | Stochastic, repel only | (0, 1) | `d`, `bw`, `coupling` |
| `"cross"` | Stochastic, with crossings | unbounded | `d`, `bw`, `coupling` |

In all cases `d` sets the Euclidean distance between the two output series,
and `w` is returned as the ground-truth coupling weight. A custom `w` vector
or function `f(n)` can also be passed directly to `sim_trends()`.
